Sequential rank agreement methods for comparison of lists

### 2015-08-07 latex/paper.Rnw
fine tuning of application and discussion sections

Incomplete lists are also a common occurrence. They arise for example
in case of missing data (items), when comparing top $d$ list results
from publications, or when some methods only rank a subset of the
items. For example, penalized regression based on the Lasso provides a
sparse set of predictors that have non-zero coefficients. There is no
obvious ordering of the set of predictors whose coefficient has been
shrunk to zero. Incomplete lists also occur when the analyst decides
to censor all lists at depth $d$ such that the measurement of
agreement is restricted to the first $d$ elements of each list.

@@

\section{Application}
We now consider an application of the sequential rank agreement to two
data sets consisting of MALDI-TOF (Matrix-Assisted Laser
Desorption/Ionization Time Of Flight) mass spectra obtained from blood
samples from patients with either benign or malignant ovarian
tumors. The data sets are subsamples of the Danish MALOVA and DACOVA
study populations.

The MALOVA study is a multidisciplinary Danish study on ovarian cancer
\citep{Hogdall:2004:Cancer:15160342} where all Danish women diagnosed
with an ovarian tumor and referred for surgery from the participating
departments of gynecology were enrolled continuously from December
1994 to May 1999. For the purpose of illustration we use a random
subsample of $119$ patients with a total of $58$ patients with
malignant ovarian cancers as cases and $61$ patients with benign
ovarian tumors as controls. The DACOVA study is another
multidisciplinary Danish study on ovarian cancer which included about
$66\%$ of the female population of Denmark
\citep{bertelsen1991protocol}. The study aimed to continuously enroll
all patients that were referred to surgery of an ovarian tumor
clinically suspected to be cancer during the period from 1984 to
1990. Similarly, we use a random subsample from the DACOVA study of
$113$ patients with a total of $54$ malignant ovarian cancers and $59$
benign ovarian tumors/gynecologic disorders.

Each spectrum consists of $49642$ samples over a mass-to-charge range of $800$ to $20000$ which we down-sample on an equidistant grid of 5000 points by linear interpolation. We preprocess each spectrum by first removing the slow-varying baseline intensity with the SNIP algorithm \citep{ryan1988snip} followed by a normalization with respect to the total ion count. We then standardize the 5000 predictors to have column-wise zero mean and unit variance in each data set.

We use the two data sets to illustrate how the sequential rank agreement can be applied in two different scenarios. In the first scenario we assess the agreement of four different statistical classification methods in how they rank the predictors according to their importance with respect to the binary outcome. In the second scenario we assess the agreement among rankings of individual predicted risk probabilities of having a malignant tumor using the four methods. The first scenario is relevant in the context of biomarker discovery and the latter is important when e.g., ranking patients according to immediacy of treatment.

Four classification methods are considered: Random Forest
\citep{breiman2001random} implemented in the R package
\texttt{randomForest} \citep{liaw2002classification}, logistic Lasso
\citep{tibshirani1996regression} and Ridge regression
\citep{segerstedt1992ordinary} both implemented in the R package
\texttt{glmnet} \citep{friedman2010regularization}, and Partial Least
Squares Discriminant Analysis (PLS-DA) \citep{boulesteix2004pls}
implemented in the R package \texttt{caret} \citep{Jed-Wing:2014aa}.

In both scenarios we use the MALOVA data to train the statistical
models, and
%the two types of
in both situations the
agreements are assessed with respect to
perturbations of the training data in the following manner. We
repeatedly draw a sub-sample (without replication) consisting of 90\%
of the MALOVA observations and train the four models on the
sub-sample. Each of the models produces a ranking of the 5000
predictors according to their importance with respect to
discriminating between the tumor types, and the agreement of these
rankings across the different sub-samples are used to assess the
stability of ranking the importance of the predictors. For the Random
Forest classifier the predictors are ranked according to the Gini
index, while for the logistic Lasso and Ridge regression models we
order the predictors by absolute magnitude of the estimated regression
coefficients. For the PLS-DA model the importance of the predictors is
based on a weighted sum of the absolute coefficients where the weights
are proportional to the reduction in the sums of squares across the
components. To assess the stability of the individual risk predictions
we apply the predictors from the DACOVA data set to each of the
models. These predicted probabilities are then ranked in decreasing
order, and the agreement of the risk predictions are similarly
obtained by the sequential rank agreement applied across the different
sub-samples of the training data. We use 1000 iterations for the
sub-sampling procedure.

All four models depend on tuning parameters. We estimate these by a 20
times repeated 5-fold cross-validation on the training data in each
iteration. The implementation of Lasso and Ridge regression in the
\texttt{glmnet} package offers three different cross-validated
optimization criteria for the penalty parameter: total deviance,
classification accuracy and area under ROC. We apply all three
criteria to our data to investigate their effect on the agreement.

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=.5\textwidth]{pics/riskAgreementPlot}%
\includegraphics[width=.5\textwidth]{pics/predictorAgreementPlot}
\end{center}
 \caption{Sequential rank agreement of predicted risk probabilities for the DACOVA data (left) and predictor importance (right) based on 1000 sub-samples of the MALOVA training data.}
 \label{fig:app1}
\end{figure}

Figure \ref{fig:app1} (left) shows the sequential rank agreement
applied to the estimated risk probabilities of the $113$ patients in
the DACOVA data set obtained from the $1000$ sub-samples of the MALOVA
data, and Figure \ref{fig:app1} (right) shows the sequential rank
agreement of the estimated importance of the 5000 predictors. For
clarity we only show the agreement for the latter up to list depth 600
where all the curves have approximately become stable.  The Lasso
models produce censored lists depending on the value of the penalty
parameter. Under the deviance optimization criterion the median number
of non-zero coefficients was 33 (range 18 to 52). In order to
calculate the sequential rank agreement we therefore applied Algorithm
\ref{sra-algorithm} with $B = 1000$ and $P=5000$. The effect of the
censoring caused by the sparsity of the Lasso is clearly visible in
the agreement curves.

For most of the sequential rank agreement curves we see, as expected, that they start low followed by an increase until they approximately become constant.  This has the interpretation that the classification methods show higher agreement with respect to the highest ranked items (the most high-risk patients and the most important predictors respectively) than for the items deeper in the lists across the different sub-samples. The change-points where the curves become approximately constant are the list depths where the ranks of the remaining items become close to random.

The only exceptions to the expected shape of the agreement curves are the predictor importance agreements from the Ridge models and the risk prediction agreement from the Lasso model with the AUC criterion. They all show higher disagreement in the top of the lists followed by a decrease. The explanations for their behavior is different for the two methods. For the Lasso model the reason is that optimizing the penalty parameter with respect to the AUC criterion tends to favor a very high penalty value causing only a single predictor to be selected in each of the 1000 iterations. This results in a lack of generalizability to the DACOVA data which gives rise to the higher disagreement in the predicted risk probabilities. In the extreme case where the penalty becomes so high that none of the predictors are selected by the Lasso, the sequential rank agreement for the predicted probabilities becomes undefined since all the ranks will be ties.

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=.5\textwidth]{pics/ridgeCensoring}%
\includegraphics[width=.5\textwidth]{pics/aucPlot}
\end{center}
 \caption{Sequential rank agreement for Ridge regression with an artificial censoring on the absolute value of the regression coefficients (left) and box plots of AUC values across the $1000$ sub-samples with respect to the known class labels of the DACOVA data (right).}
 \label{fig:app2}
\end{figure}

The reason behind the behavior of the sequential rank agreement of predictor importances estimated by the Ridge models is more subtle. Looking at the distribution of the regression coefficients it is seen that a large proportion of them are numerically very close to zero and of almost equal absolute value. This implies that when they are ranked according to their magnitude, their actual order becomes more uncertain and more close to a random permutation. This problem can be alleviate by introducing an artificial censoring depending on the absolute value of the coefficients. Figure \ref{fig:app2} (left) shows the sequential rank agreement curve of the rankings of the predictor importances from the Ridge model using the deviance criterion (same curve as in figure \ref{fig:app1} (left)) as well as an agreement curve where the absolute value of the regression coefficients have been censored at the $0.1\%$ quantile. In our application the median number of censored predictors across the 1000 sub-samples is 5 (range 1 to 11). Even though the number of censored predictors is very small compared to the total number of predictors, the effect on the sequential rank agreement is substantial as it has now flipped to show higher agreement in the top of the lists as was expected.

Returning to Figure \ref{fig:app1}, it can further be seen that some of the methods show better agreement with respect to the predicted probabilities than for ranking the importance of the predictors and vice versa. Ridge regression performs very well for the risk predictions, but PLS-DA performs better for the predictor importances and similarly worse for the risk predictions. The Random Forest method seems to perform equally well in both scenarios. Lasso regression (except with the AUC criterion) performs very well at ranking the high-risk patients but very poorly at ranking the importance of the predictors. This reason for this is the high auto-correlation between the intensities in the mass spectra which leads to collinearity issues in the regression models. It is well-known that variable selection with the Lasso penalty does not preform very well when the predictors are highly correlated. This collinearity does, however, not affect the agreement of the risk predictions, as seen on the left part of the figure, since it is not so important which specific variable that gets selected from a group of highly correlated predictors when the purpose is risk predictions.

Finally, it would appear that Ridge regression with the AUC criterion for the penalty parameter achieves the best performance with respect to the stability of ranking the individual predicted risk probabilities. It must, however, be stressed that the sequential rank agreement in this application is only concerned with whether similar orderings of the patients is obtained under perturbations of the training data. It is not concerned with whether such an ordering is the correct one. This is similar to the usual distinction between accuracy and precision. Looking at the distributions of the AUCs illustrated in Figure \ref{fig:app2} (left) across the $1000$ sub-samples for a selection of the models with respect to the known class labels of the DACOVA data, we see that PLS-DA attains the highest AUC with a median value of $0.71$ while the Ridge model with the AUC criterion attains a median AUC of $0.49$. This implies that while Ridge regression optimized with respect to the AUC criterion achieves the best sequential rank agreement it performs similar to a random coin toss with respect to classifying the DACOVA patients. Both concerns are of course important in practice and for this data set it appears that PLS-DA is the overall best method taking both agreement and classification accuracy into account.

\section{Discussion}
In this article we address the problem of comparing ranked lists of
the same items. Our proposed method can handle both the situation
where the underlying data to generate the ranked lists are available
and the situation where the only available data is the actual ranked
lists. In addition, censored ranked lists where only the top $k$
ranked items are known can be accommodated as well. The proposed
agreement measure can be interpreted as the average distance between
an item's rank and the average rank assigned to that item and thus
determine how well the lists agree on the ranks of a specific
item and it circumvents the problem with other existing similarity measures.

The sequential rank agreement is extremely versatile. We have shown
that it can be used not only to compare ranked lists of items produced
from different samples/populations but that it also can be used to
study the ranks obtained from different analysis methods on the same
data as well as to evaluate the stability of the ranks from a single
method by bootstrapping the data repeatedly and comparing the ranks
obtained from the bootstrapped data.

@@

Finally, we have --- whenever possible --- used as much information
from the lists as possible, i.e., we have used all available ranks.
We could choose to censor the rankings of all the items that were
deemed not to be significant if information on both rankings and the
corresponding evidence for significance was available. That would
ensure that there would be put less emphasis on the agreement of the
non-significant items and it would be easier to identify a change in
agreement among the items that were deemed to relevant.

We note that the sequential rank agreement is still marred by problems
that generally apply to ranking of items and/or
individuals. Collinearity in particular can be a huge problem when
bootstrapping data or when comparing different analysis methods. For
example, marginal analyses where each item is analyzed separately will
assign similar ranks to two highly correlated predictors while methods
that provide a sparse solution such as the lasso will just rank one of
the two predictors high while the other might have a very low
rank. When the rankings from these two analysis methods are compared
then the agreement will differ simply because of the way correlated
predictors are handled. This is not a shortcoming of the sequential
rank agreement but is a problem general to all ranked lists.

Another caveat with the way the sequential rank agreement is defined
is the use of the standard deviation to measure agreement. The
standard deviation is an integral part of the limits-of-agreement as
discussed by \citet{alt:bland:1983}. However, the standard deviation
can also be unstable when the number of observations is low and
alternative measures such as the mean absolute deviance may prove more
stable in some situations. However, the current definition using the
standard deviation is analogous to the approach used for agreement in
method comparisong studies.

In conclusion we have introduced a method for evaluation of ranked
(censored) lists that can be easily interpreted and that can be
applied to a large number of situations.  The method can be adapted
even further by using it to identify and cluster statistical analysis
methods that provide more or less the same agreement, or by using the
rank agreement to optimize a hyper-parameter in, say, elastic net regularized
regression where the rank agreement is used to determine the mixing
proportion between the $L_1$ and the $L_2$ penalty. Finally, the
proposed method may be adapted (with some additional assumptions) to
the situation where the lists are subsets of each other. We will be
investigating these extensions further in the future.

>>>>

Incomplete lists are also a common occurrence. They arise for example
in case of missing data (items), when comparing top $d$ list results
from publications, or when some methods only rank a subset of the
items. For example, penalized regression based on the LASSO provides a
sparse set of predictors that have non-zero coefficients. There is no
obvious ordering of the set of predictors whose coefficient has been
shrunk to zero. Incomplete lists also occur when the analyst decides
to censor all lists at depth $d$ such that the measurement of
agreement is restricted to the first $d$ elements of each list.

@@

\section{Application}
We now consider an application of the sequential rank agreement to two
data sets consisting of MALDI-TOF (Matrix-Assisted Laser
Desorption/Ionization Time Of Flight) mass spectra obtained from blood
samples from patients with either benign or malignant ovarian
tumors. The data sets are sub-samples of the Danish MALOVA and DACOVA
study populations.

The MALOVA study is a multidisciplinary Danish study on ovarian cancer
\citep{Hogdall:2004:Cancer:15160342} where all Danish women diagnosed
with an ovarian tumor and referred for surgery from the participating
departments of gynecology were enrolled continuously from December
1994 to May 1999. For the purpose of illustration we use a random
sub-sample of $119$ patients with a total of $58$ patients with
malignant ovarian cancers as cases and $61$ patients with benign
ovarian tumors as controls. The DACOVA study is another
multidisciplinary Danish study on ovarian cancer which included about
$66\%$ of the female population of Denmark
\citep{bertelsen1991protocol}. The study aimed to continuously enroll
all patients that were referred to surgery of an ovarian tumor
clinically suspected to be cancer during the period from 1984 to
1990. Similarly, we use a random sub-sample from the DACOVA study of
$113$ patients with a total of $54$ malignant ovarian cancers and $59$
benign ovarian tumors/gynecologic disorders.

Each spectrum consists of $49642$ samples over a mass-to-charge range
of $800$ to $20000$ which we down-sample on an equidistant grid of
5000 points by linear interpolation. After this dimension reduction
the mass spectrum is represented by 5000 predictors. We preprocess
%% ANDREAS: det er uklar om preprocessing forgaar i de oprindelig spectra eller i de down-samplede 
each spectrum by first removing the slow-varying baseline intensity
with the SNIP algorithm \citep{ryan1988snip} followed by a
normalization with respect to the total ion count. We then standardize
the 5000 predictors to have column-wise zero mean and unit variance in
each data set.

We use the two data sets to illustrate how the sequential rank
agreement can be applied in two different scenarios. In the first
scenario we assess the agreement of four different statistical
classification methods in how they rank the predictors according to
their importance for distinguishing benign and malignant tumors. In
the second scenario we assess the agreement among rankings of
individual predicted risks of having a malignant tumor. The first
scenario is relevant in the context of biomarker discovery and the
latter is important, e.g., to rank patients according to immediacy of
treatment.

Four classification methods are considered: Random Forest
\citep{breiman2001random} implemented in the R package
\texttt{randomForest} \citep{liaw2002classification}, logistic LASSO
\citep{tibshirani1996regression} and Ridge regression
\citep{segerstedt1992ordinary} both implemented in the R package
\texttt{glmnet} \citep{friedman2010regularization}, and Partial Least
Squares Discriminant Analysis (PLS-DA) \citep{boulesteix2004pls}
implemented in the R package \texttt{caret} \citep{Jed-Wing:2014aa}.

In both scenarios we use the MALOVA data to train the statistical
models, and in both situations the agreements are assessed with
respect to perturbations of the training data in the following
manner. We repeatedly draw a random sub-sample (without replication)
consisting of 90\% of the MALOVA observations and train the four
models on each sub-sample. We use 1000 iterations for the sub-sampling
procedure.

\subsection{Agreement of predictor rankings}

For each of the four methods, each of the 1000 models trained in on of
the 1000 sub-samples of the MALOVA data produces a ranking of the 5000
predictors according to their importance for discriminating between
the tumor types.

All four methods depend on tuning parameters. We select these
separately for each sub-sample by a 20 times repeated 5-fold
cross-validation.
%% ANDREAS: hvilke tuning parameters har du optimeret for Random forest og PLS-DA?
The implementation of LASSO and Ridge regression in the
\texttt{glmnet} package offers three different cross-validated
optimization criteria for the penalty parameter: total deviance,
classification accuracy and area under ROC. We apply all three
criteria to our data to investigate their effect on the agreement.
Note also that the LASSO models produce censored lists depending on
the value of the penalty parameter. Under the deviance optimization
criterion the median number of non-zero coefficients was 33 (range 18
to 52). In order to calculate the sequential rank agreement for the
LASSO we therefore applied our proposed Algorithm \ref{sra-algorithm}
with $B = 1000$ and $P=5000$.

For the Random Forest classifier the predictors are ranked according
to the Gini index, while for the logistic LASSO and Ridge regression
models we order the predictors by absolute magnitude of the estimated
regression coefficients. For the PLS-DA model the importance of the
predictors is based on a weighted sum of the absolute coefficients
where the weights are proportional to the reduction in the sums of
squares across the components.

The right panel of Figure \ref{fig:app1} shows the sequential rank
agreement of the estimated importance of the 5000 predictors.  For
clarity of presentation we zoom in on the agreement up to list depth
120. At deeper list depths all agreement curves appeared stable.

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=.5\textwidth]{pics/riskAgreementPlot}%
\includegraphics[width=.5\textwidth]{pics/predictorAgreementPlot}
\end{center}
%% ANDREAS: kan du aendre titel for begge grafer:
%% "Risk probabilities" -> "Predicted risks of malignant tumor in 113 patients (DACOVA)"
%% "Predictor importance" -> "Discrimination importance of 5000 predictors"
\caption{Left panel: Sequential rank agreement of 1000 rankings of the
  predicted risks of malignant tumor. For each method the different
  rankings were obtained by first training models in 1000 random
  sub-samples of the MALOVA data and then predicting the risk of
  malignant tumor in the 113 MALOVA patients. Right panel: Sequential
  rank agreement of 1000 rankings of the 5000 predictors. The rankings
  were obtained from the same 1000 trained models.}
 \label{fig:app1}
\end{figure}

For most of the sequential rank agreement curves we see, as expected,
that they start low, indicating good agreement, followed by an
increase until they approximately become constant. This has the
interpretation that the agreement across the different sub-samples is
higher in the top as compared to the tail of the lists for all these
classification methods. The change-points where the curves become
approximately constant are the list depths where the ranks of the
remaining items become close to random.
%% ANDREAS: boer "close to random" ikke vaere ens for alle metoder?

A not expected shape of the agreement curves is seen for the Ridge
models for all three tuning criteria. They all show higher
disagreement in the top of the lists followed by a decrease. The
reason behind this behavior is rather subtle. Looking at the
distribution of the regression coefficients we see that a large
proportion of them are numerically very close to zero and have almost
equal absolute value. This is a general feature of the Ridge models in
this data set and seen for all the 1000 trained models.
%% ANDREAS: er det sidste ok?
This implies that when predictors are ranked according to the
magnitude of their coefficients, their actual order becomes more
uncertain and more close to a random permutation. This problem can be
alleviated by artificially censoring all predictors with absolute
coefficient values below a given threshold. For the Rigde models tuned
with the deviance criterion, Figure \ref{fig:app2} (left) shows the
the sequential rank agreement where for each of the 1000 trained
models predictors were artificially censored when their absolute
coefficient value was lower than the $0.1\%$ quantile of the 5000
absolute coefficient values. The corresponding curve from Figure
\ref{fig:app1} (left panel) is shown for comparison.
%% ANDREAS: godkender du min aendring af de sidste 2 saetninger?
Thus, in our application the median number of censored predictors
across the 1000 sub-samples was 5 (range 1 to 11). Even though the
number of censored predictors is very small compared to the total
number of predictors, the effect on the sequential rank agreement is
substantial and with the artificial censoring the shape of the curves
is as expected, starting low and then increasing. Also, the effect of
the censoring caused by the sparsity of the LASSO is clearly visible
in the agreement curves.
%% ANDREAS: man kan desvaerre ikke se hvor LASSO agreement forsvinder
%% (ved list-depth 5?) og om den er laver for list depth 1-5 end rigde
%% eller om den er kun laver for list depth 1-2. maaske kan du
%% beskrive dette i teksten?

\subsection{Agreement of individual risk predictions}
\label{sec:airp}

To assess the stability of the individual risk predictions we apply
the predictors from the DACOVA data set to each of the models. The
predicted probabilities are then ranked in decreasing order such the
patients with the highest risk of a malignant tumor appears on the top
of the list. Figure \ref{fig:app1} (left) shows the sequential rank
agreement separately for each method, based on 1000 risk predictions
obtained from the models trained in the same $1000$ random sub-samples
of the MALOVA data. For clarity of presentation we zoom in on the
agreement up to list depth 600. At deeper list depths all agreement
curves appeared stable.

Most curves start low and then increase indicating higher agreement
among high risk patients.
%% ANDREAS og CLAUS: er det ogsaa forvented eller skyldes det at vi kumulerer fra toppen?
%% agreement boer jo igen vaere hoejt bland low risk patienter ...
An exception is the risk prediction agreement for the LASSO tuned with
the AUC criterion which shows very low agreement among the high values
of the predicted risks. The reason is that optimizing the penalty
parameter with respect to the AUC criterion tends to favor a very high
penalty value causing only a single predictor to be selected in each
of the 1000 iterations. This results in a lack of generalizability to
the DACOVA data which gives rise to the higher disagreement in the
predicted risk probabilities. In the extreme case where the penalty
becomes so high that none of the predictors are selected by the LASSO,
the sequential rank agreement for the predicted probabilities becomes
undefined since all the ranks will be ties.

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=.5\textwidth]{pics/ridgeCensoring}%
\includegraphics[width=.5\textwidth]{pics/aucPlot}
\end{center}
\caption{Left panel: Sequential rank agreement for Ridge regression
  obtained by artificially censoring predictors when their absolute
  coefficient values are lower than the 0.1\% quantile. Right panel:
  Box plots of AUC values across the $1000$ sub-samples with respect
  to the known class labels of the DACOVA data.}
 \label{fig:app2}
\end{figure}

In Figure \ref{fig:app1} it can further be seen that some of the
methods show better agreement with respect to the predicted
probabilities than for ranking the importance of the predictors and
vice versa. Ridge regression shows higher agreement across training
sets for the risk predictions than PLS-DA, and PLS-DA shows higher
agreement for predictor importance than Ridge regression. The Random
Forest method seems to perform equally well in both scenarios.
%% Andreas: jeg forstaar ikke den sidste saetning. kan man sammenligne de to rank agreements (left vs right panel)?
LASSO shows similar agreement for ranking the risk predictions than
PLS-DA (unless with the AUC criterion), and poor agreement for ranking
predictors. This reason for the latter is the high auto-correlation
between the intensities in the mass spectra which leads to
collinearity issues in the regression models. It is well-known that
variable selection with the LASSO penalty does not preform very well
when the predictors are highly correlated. The collinearity does,
however, not affect the agreement of the risk predictions, as seen on
the left panel of Figure \ref{fig:app1}, since it is not so important
which specific variable is selected from a group of highly correlated
predictors when the purpose is risk predictions.

It appears that Ridge regression tuned with the AUC criterion achieves
the best performance with respect to the stability of ranking the
individual predicted risk probabilities. It must, however, be stressed
that the sequential rank agreement in this application is only
concerned with the agreement of the risk predictions across
sub-samples and not with the actual accuracy of the risk
predictions. Thus, we also computed the AUC for the different models
based on the DACOVA data. The distributions across the $1000$
sub-samples for a selection of the models is shown in the right panel
of Figure \ref{fig:app2}.  Here we see that PLS-DA attains the highest
AUC values with a median value of $0.71$ while the Ridge model with
the AUC criterion attains a median AUC of $0.49$. This implies that
while Ridge regression optimized with respect to the AUC criterion
achieves the best sequential rank agreement it performs similar to a
random coin toss with respect to classifying the DACOVA patients.
%% and for this data set it appears that PLS-DA is the overall best
%% method taking both agreement and discrimination ability into account.
%% Andreas: vi boer ogsaa se paa Brier score (calibration) inden vi konkluderer at PLS-DA er den bedste

\section{Discussion}

In this article we address the problem of comparing ranked lists of
the same items. Our proposed method can handle both the situation
where the underlying data to generate the ranked lists are available
and the situation where the only available data is the actual ranked
lists. In addition, censored ranked lists where only the ranks of the
top $k$ ranked items are known can be accommodated as well. The
proposed agreement measure can be interpreted as the average distance
between an item's rank and the average rank assigned to that item
across lists. Thus the measure determines how well the lists agree on
the ranks of a specific item.

The sequential rank agreement is extremely versatile. We have shown
that it can be used not only to compare ranked lists of items produced
from different samples/populations but that it also can be used to
study the ranks obtained from different analysis methods on the same
data as well as to evaluate the stability of the ranks from a single
method by bootstrapping (or sub-sampling) the data repeatedly and
comparing the ranks obtained from training the models in the
bootstrapped data.

@@

Finally, we have --- whenever possible --- used as much information
from the lists as possible, i.e., we have used all available ranks.
%% Claus: hvad vil du sige med den sidste saetning?
We could choose to censor the rankings of all the items that were
deemed not to be significant if information on both rankings and the
corresponding evidence for significance was available. That would
ensure that there would be put less emphasis on the agreement of the
non-significant items and it would be easier to identify a change in
agreement among the items that were deemed to relevant. In our
application section we have successfully introduced such an artificial
censoring for the predictor rankings obtained with ridge regression.

We note that the sequential rank agreement is still marred by problems
that generally apply to ranking of items and/or
individuals. Collinearity in particular can be a huge problem when
bootstrapping data or when comparing different analysis methods. For
example, marginal analyses where each item is analyzed separately will
assign similar ranks to two highly correlated predictors while methods
that provide a sparse solution such as the LASSO will just rank one of
the two predictors high while the other might have a very low rank.
Thus in such a scenario we would expect low agreement of the rankings
from LASSO and marginal analyses simply because of the way correlated
predictors are handled. This is not a shortcoming of the sequential
rank agreement but is a problem general to all ranked lists.

Another caveat with the way the sequential rank agreement is defined
is the use of the standard deviation to measure agreement. The
standard deviation is an integral part of the limits-of-agreement as
discussed by \citet{alt:bland:1983}. However, the standard deviation
can also be unstable when the number of observations is low and
alternative measures such as the mean absolute deviance may prove more
stable in some situations. However, the current definition using the
standard deviation is analogous to the approach used for agreement in
method comparison studies.

%% In conclusion we have introduced a method for evaluation of ranked
%% (censored) lists that can be easily interpreted and that can be
%% applied to a large number of situations.
%% CLAUS: den sidste saetning er bar en omformulering af Diskussionen's foerste saetning.

%% The method here proposed can be adapted further by using it to
%% identify and cluster statistical analysis methods that provide more or
%% less the same agreement,
%% Claus: det er enten for kort eller for lang

%% or by using the rank agreement to optimize a
%% hyper-parameter in, say, elastic net regularized regression where the
%% rank agreement is used to determine the mixing proportion between the
%% $L_1$ and the $L_2$ penalty.
%% Claus: det er enten for kort eller for lang

%% Finally, the proposed method may be
%% adapted (with some additional assumptions) to the situation where the
%% lists are subsets of each other.
%% Claus: det er enten for kort eller for lang

%% We will be investigating these
%% extensions further in the future.
